Exploring the Motives and Driving Forces Behind Modifications to APIs From a Software Evolution Perspective

### 2015-05-07 thesis/paper.tex
Updated related work etc

%Structure of the paper
The sections following the Introduction are structured as follows: In Section \ref{existing_literature}, we present the background and previously published work that is related to our study. Section \ref{case_company_description} describes the case company and how it was selected. A description of our methodology is introduced in Section \ref{methodology}. In Section \ref{results}, the results of our study are presented and in Section \ref{conclusion}, we discuss the findings of our study.

@@

\subsection{API Design} \label{api_design}
API design is notoriously difficult, as a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating} \cite{bloch2008effective} \cite{stylos2006comparing}. Examples of such decisions are whether to use the factory pattern or constructors to create objects, or if the API should display errors during compilation or runtime \cite{stylos2006comparing}. More trivial design problems, such as, assigning names to API features and difficulty in naming types also have a large impact on the usability of the API \cite{shi2011empirical}. When facing such design decisions, the following four factors are important to consider: a) The API  must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API is clear to the user when, for example, using design patterns, such as the concrete factory pattern \cite{stylos2006comparing} \cite{shi2011empirical}. The sum of all design decisions taken during the development of an API will decide the usability of the API. Measuring such a factor can be done by measuring the twelve cognitive dimensions which are factors that are impacted by the interaction between the API and the user of it \cite{clarke2004measuring}.

\subsection{Software Evolution} \label{software_evolution}
%%Categories of maintenance and types of maintenance
Software evolution is a field which studies the application of software maintenance activities and processes on software and the resulting, evolved versions of the software. The concept of software maintenance has existed since the 1960s. Well-known studies from this era established the importance of software maintenance \cite{lientz1980software}. One study would go on to have a particularly profound effect on software maintenance by establishing a set of four categories which describe the intents behind change to software \cite{lientz1980software}. These categories were refined in a further study, in which the number of categories were increased from four to twelve \cite{chapin2001types}. The categories proposed were: Enhancive, corrective, reductive, adaptive, performance, preventive, groomative, updative, reformative, evaluative, consultive and training \cite{chapin2001types}. Each category describes a type of software maintenance activity which relate to one of three particular areas; the code, the software and the customer-experienced functionality \cite{chapin2001types}. This study also helped to further the distinction between software maintenance and software evolution. The study suggested that software maintenance is a part of software evolution and that the two are not interchangeable \cite{chapin2001types}.

%%Lehman's laws (!!!improve later)
An early study made in the area of software evolution suggests that changes to software must obey one of eight laws of software evolution \cite{lehman1980programs}. The laws presented in the study are 1) Continuing Change, 2) Increasing Complexity, 3) Self Regulation, 4) Conservation of Organisational Stability, 5) Conservation of Familiarities, 6) Continuing Growth, 7) Declining Quality and 8) Feedback System \cite{lehman1980programs}. The study also outlines that these laws only apply to E-programs, which are programs that are ``change prone" and that ``mechanise a human or societal activity" \cite{lehman1980programs}.

\subsection{Related Work} \label{related_work}
Previous studies that have attempted analyse APIs in the context of software evolution have primarily focused on APIs that are a part of large programming languages, such as Java \cite{hou2011exploring} \cite{shi2011empirical} and Smalltalk \cite{robbes2012developers}. One study \cite{hou2011exploring} attempted to uncover the intents behind the changes made in the AWT and Swing Java libraries. They found that the use of a strong architecture was vital in ensuring a successful evolution of the API \cite{hou2011exploring}. In their study, the intents were categorised around programming related factors such as reducing coupling or conformance to naming conventions.

One study \cite{dig2005role} investigated three frameworks and one library in which the study found that 80\% of refactoring changes to APIs negatively affected existing applications. Refactorings are not the only changes to APIs that can impact existing projects negatively. Another study was conducted regarding the changes, which the researchers refer to as ripple effects \cite{robbes2012developers}. They found that changes in APIs have an effect upon the entire ecosystem. This particular empirical study found that 14\% of non-trivial API deprecations caused errors in at least one project, with the worst case of 79 projects being affected \cite{robbes2012developers}.

>>>>

%Structure of the paper
The sections following the Introduction are structured thusly: In Section \ref{existing_literature}, we present the background and previously published work that is related to our study. Section \ref{case_company_description} describes the case company and how it was selected. A description of our methodology is introduced in Section \ref{methodology}. In Section \ref{results}, the results of our study are presented and in Section \ref{conclusion}, we discuss the findings of our study.

@@

\subsection{API Design} \label{api_design}
API design is notoriously difficult, a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating} \cite{bloch2008effective} \cite{stylos2006comparing}. Examples of such decisions include whether or not to make use of the factory pattern or if an API should display errors only at compilation or only at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design decisions, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API is clear to the user, for example, using design patterns \cite{stylos2006comparing} \cite{shi2011empirical}. The design decisions reached during development of an API will effect the overall usability of the API, for better or worse. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and the API's user \cite{clarke2004measuring}.

\subsection{Software Evolution} \label{software_evolution}
%%Categories of maintenance and types of maintenance
Software evolution is a field that studies the application of software maintenance activities, changes in software processes and the resulting, evolved versions of the software. The concept of software maintenance has existed since the 1960s when it was first introduced to the software development community \cite{lientz1980software}. A set of four categories describing software maintenance categories \cite{lientz1980software} became the basis upon which twelve new categories were created \cite{chapin2001types}. The tweleve categories are: enhancive, corrective, reductive, adaptive, performance, preventive, groomative, updative, reformative, evaluative, consultive and training \cite{chapin2001types}. Each category describes a software maintenance activity that relates to one of three particular areas; the code, the software and the customer-experienced functionality \cite{chapin2001types}. It has been suggested that software maintenance is a part of software evolution and that the two terms are not interchangeable \cite{chapin2001types}.

%%Lehman's laws (!!!improve later)
Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman1980programs}. The laws, known as Lehman's laws, are: 1) Continuing Change, 2) Increasing Complexity, 3) Self Regulation, 4) Conservation of Organisational Stability, 5) Conservation of Familiarities, 6) Continuing Growth, 7) Declining Quality and 8) Feedback System \cite{lehman1980programs}. The laws are said to only apply to E-type systems, which are programs that are ``change prone" and that ``mechanise a human or societal activity" \cite{lehman1980programs}.

\subsection{Related Work} \label{related_work}
The analysis of APIs in the context of software evolution has primarily focused on APIs that are a part of large programming languages, such as Java \cite{hou2011exploring} \cite{shi2011empirical} and Smalltalk \cite{robbes2012developers}. Attempts to uncover the intents behind the changes made in the AWT and Swing Java libraries have also been undertaken \cite{hou2011exploring}, leading to suggestions that the use of a strong architecture is vital for ensuring the successful evolution of an API \cite{hou2011exploring}.

Investigations of three frameworks and one library indicated that 80\% of refactoring changes to APIs negatively affected existing applications \cite{dig2005role}. The changes, sometimes referred to as ripple effects \cite{robbes2012developers} indicate that negative effects of changes can propagate throughout a software ecosystem. It was found that 14\% of non-trivial API deprecations caused errors in at least one project, with the worst case of 79 projects being affected \cite{robbes2012developers}.

%%Possibly add a paragraph here to
